Asimilación de distintas observaciones

### 2020-12-18 analysis/chapters/experimento_20181122.Rmd
Post reunión 18/12

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")

derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"

@@

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))

@@

```
```{r rad-wv}
files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles/asim_iasi*ensmean"))

rad <- read_diag_rad(files, "E7") %>% 
  .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
  .[wavenumber %between% c(1367, 1422.25)] %>% 
  satinfo[., on = c("sensor", "channel")]

rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel %between% c(330, 335) & 
      qc == 0 & iuse == 1] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(trans = mesoda::symlog_trans(),
                        guide = guide_colourbar(barheight = 18)) +
  geom_mapa() +
  facet_wrap(~date) +
  labs(x = NULL, y = NULL, color = NULL,
       caption = "1410.75 cm^-1") +
  theme_minimal()
```

@@

```{r}
files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")

sondeos %>% 
  .[variable == "t" & !str_detect(site, "/") & 
      exp == "E4"] %>% 
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = value), size = 0.5) +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  # geom_contour(aes(z = value - fcst_value)) +
  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  labs(color = "T", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  sondeos %>% 
  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
  .[, oma := value - fcst_value] %>% 
  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = oma), size = 0.5) +
  scale_color_divergent() +
  # geom_contour(aes(z = value - fcst_value)) +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                               E6 = "SATWND", E7 = "RAD"))) +
  labs(color = "OmA", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

```{r}
sondeos %>% 
  .[variable == "td" & !str_detect(site, "/") & 
      exp == "E4"] %>% 
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = value), size = 0.5) +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  # geom_contour(aes(z = value - fcst_value)) +
  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  labs(color = "Td", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

@@

#### FSS for the forecast

>>>>

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"

@@

@@

```
```{r rad-wv}
files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles*/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles*/asim_iasi_metop-b*ensmean"))

rad <- purrr::map(files, function(f) {

  rad <- read_diag_rad(f, "E7") %>% 
    .[, run := basename(dirname(f))] %>% 
    .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
    .[wavenumber %between% c(1367, 1422.25)] %>% 
    satinfo[., on = c("sensor", "channel")]
}) %>% rbindlist()

rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      qc == 0 & iuse == 1] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(trans = mesoda::symlog_trans(),
                        guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
                                                   "diagfiles_ana" = "OmA"))) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal() +

  rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      # qc == 0 & 
      iuse == 1] %>% 
  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
  .[, ":="(tbc = diagfiles - diagfiles_ana,
           run = "AmB")] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  labs(x = NULL, y = NULL, color = NULL) +
  facet_grid(run~date) +
  theme_minimal() +
  plot_layout(ncol = 2, widths = c(0.4, 0.6))
```

@@

```{r temp}
fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
                     exp = c("E4", "E5", "E6", "E7"))

t <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, t := ReadNetCDF(paste0(derived_data, "/t_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(t = "temp"), out = "vector")] %>% 
    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

t[date == unique(date)[1:3] & west_east %between% c(70, 180)] %>% 
  .[, .(t_mean = mean(t),
        u_mean = mean(u),
        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
  .[, ":="(dir = atan2(v_mean, u_mean)*180/pi,
           mag = Mag(u_mean, v_mean))] %>%
  .[, date := factor(date)] %>% 
  ggplot(aes(south_north, bottom_top)) +
  geom_contour_fill(aes(z = t_mean, fill = stat(level_d))) +
  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
                       palette = "OrRd", direction = 1,
                       # limits = c(1, 19),
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.8)) +
  # geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
  #              color = "grey30", size = 0.3, binwidth = 2) +
  # geom_text_contour(aes(z = v_mean), color = "grey30", size = 4, stroke = 0.1, breaks = seq(-10, 6, 2)) +
  # scale_linetype(guide = NULL) +
  geom_arrow(aes(mag = mag, angle = dir),
             arrow.angle = 8, size = 0.2, skip.x = 10, skip.y = 3) +
  scale_mag() +
  facet_grid(date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                     E6 = "SATWND", E7 = "RAD"))) +
  labs(y =  "Sigma levels", x = "Grid points") +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

@@

```
#### Soundings

```{r message=FALSE, warning=FALSE}
files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")

IOP <- tribble(
  ~iop, ~ini, ~end,
  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
) %>% setDT()

sondeos %>% 
  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
  .[, oma := value - fcst_value] %>% 
  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = oma), size = 0.5) +
  scale_color_divergent() +
  # geom_contour(aes(z = value - fcst_value)) +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                                       E6 = "SATWND", E7 = "RAD"))) +
  labs(color = "OmA", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  sondeos %>% 
  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
  .[, oma := value - fcst_value] %>% 
  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = oma), size = 0.5) +
  scale_color_divergent() +
  # geom_contour(aes(z = value - fcst_value)) +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                                       E6 = "SATWND", E7 = "RAD"))) +
  labs(color = "OmA", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  plot_layout(ncol = 2, widths = c(0.5, 0.5))
```

sondeos %>% 
  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2)), 
           BIAS = mean(value - fcst_value)), by = .(iop, exp, variable, lev)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(iop~variable, scales = "free_x") +
  labs(x = NULL, y = NULL, color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r}
sondeos %>% 
  .[variable == "td" & !str_detect(site, "/") & 
      exp == "E4"] %>%, 
  ggplot(aes(launch_time, alt/1000)) +
  geom_point(aes(color = value), size = 0.5) +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  # geom_contour(aes(z = value - fcst_value)) +
  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
  labs(color = "Td", y = NULL, x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

### Forecast
